### 2015-02-06 landslide-lecture/2015/lecture.tex
Add rough draft of landslide lecture for p2 for s15.

\documentclass[xcolor=dvipsnames]{beamer}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{manfnt}
\usepackage{wasysym}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{url}
\usepackage{ulem}
\usepackage{marvosym}
\usepackage{skull}
\usepackage{proof}
\usepackage{array}
\usepackage{colortbl}
\setbeamertemplate{navigation symbols}{}

\title[Landslide]{{\bf Landslide: A New Race-Finding Tool for 15-410}}
\subtitle[]{ {\em more clever than } ``\texttt{mandelbrot}'' {\em since 2011.}}
\author[Ben Blum]{Ben Blum \texttt{(bblum@andrew.cmu.edu)}}

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2014, December 3}

\setbeamertemplate{footline}{\hspace*{.5cm}\scriptsize{\insertauthor\hspace*{50pt} \hfill\insertframenumber\hspace*{.5cm}}}

\usecolortheme{seahorse}
\usecolortheme{rose}
\useoutertheme{infolines}

\usecolortheme[named=ForestGreen]{structure}

\newcommand\noob{\mathsf{noob}}
\newcommand\gibs{\mathsf{gibs}}
\newcommand\dps{\mathsf{dps}}
\newcommand\squig\rightsquigarrow
\newcommand\Coloneqq{\mathrel{\mathop{::}}=}
\newcommand\dmg{\text{\Laserbeam}}
\newcommand\delter\delta
\newcommand\alpher\alpha
\newcommand\defnor{\text{ }|\text{ }}

\newcommand\pimp{\mathop{\supset}}
\newcommand\pand{\mathop{\wedge}}
\newcommand\por{\mathop{\vee}}
\newcommand\ptrue{\top}
\newcommand\pfalse{\bot}

\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{olivegreen}{RGB}{0,127,0}

\begin{document}
\renewcommand{\inserttotalframenumber}{41}
\normalem
\begin{frame}
	\titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\linegap{\vspace{0.2in}}
\newcommand\breakslide[1]{\begin{frame}{} \begin{center} #1 \end{center} \end{frame}}

\section{Introduction}
\subsection{Introduction}

\begin{frame}{Outline}
	\textbf{Theory: Seeing race conditions in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Technique: Systematic testing}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Coping with state space explosion
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

\subsection{Race Conditions}

\definecolor{thread1}{RGB}{87,172,255}
\definecolor{thread2}{RGB}{255,201,102}
\definecolor{thread3}{RGB}{255,87,130}

\begin{frame}{Case Study}
	\begin{center}
		\begin{tabular}{llll}
			\begin{tabular}{l}
				Consumer thread \\
				\hline
				\\
				\texttt{LOCK(mx);} \\
				\\
				\texttt{if~(!work\_exists())} \\
				\texttt{~~~~WAIT(cvar,~mx);} \\
				\texttt{work~=~claim\_work();} \\
				\\
				\texttt{UNLOCK(mx);} \\
				\texttt{assert(work~!=~NULL);} \\
			\end{tabular}
			& & &
			\begin{tabular}{l}
				Producer thread \\
				\hline
				\\
				\texttt{LOCK(mx);} \\
				\\
				\texttt{give\_work(work);} \\
				\texttt{SIGNAL(cvar);} \\
				\\
				\texttt{UNLOCK(mx);} \\
				\\
				\\
			\end{tabular}
		\end{tabular}
		\linegap

		\begin{itemize}
			\item See {\bf Paradise Lost} lecture!
			\item {\tt if} vs {\tt while}: Two consumers can race to make one fail.
		\end{itemize}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (``good'' case)}
	\begin{center}
		\begin{tabular}{|l|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} & \cellcolor{thread3} {\bf Thread 3} \\
			\hline
			\small \texttt{LOCK(mx);} & & \\
			\small \texttt{if~(!work\_exists())} & & \\
			\small \texttt{~~~~WAIT(cvar,~mx);} & & \\

			& \small \texttt{LOCK(mx);} & \\
			& \small \texttt{give\_work(work);} & \\
			& \small \texttt{SIGNAL(cvar);} & \\
			& \small \texttt{UNLOCK(mx);} & \\

			\small \texttt{work~=~claim\_work();~} & & \\
			\small \texttt{UNLOCK(mx);} & & \\
			\small \texttt{//~got~the~work} {\large \smiley}& & \\

			& & \small \texttt{LOCK(mx);} \\
			& & \small \texttt{if~(!work\_exists())~} \\
			& & \small \texttt{~~~~WAIT(cvar,~mx);} \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (``good'' case, again)}
	\begin{center}
		\begin{tabular}{|l|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} & \cellcolor{thread3} {\bf Thread 3} \\
			\hline
			\small \texttt{LOCK(mx);} & & \\
			\small \texttt{if~(!work\_exists())} & & \\
			\small \texttt{~~~~WAIT(cvar,~mx);} & & \\

			& & \small \texttt{LOCK(mx);} \\
			& & \small \texttt{if~(!work\_exists())~} \\
			& & \small \texttt{~~~~WAIT(cvar,~mx);} \\

			& \small \texttt{LOCK(mx);} & \\
			& \small \texttt{give\_work(work);} & \\
			& \small \texttt{SIGNAL(cvar);} & \\
			& \small \texttt{UNLOCK(mx);} & \\

			\small \texttt{work~=~claim\_work();~} & & \\
			\small \texttt{UNLOCK(mx);} & & \\
			\small \texttt{//~got~the~work} {\large \smiley}& & \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (race condition)}
	\begin{center}
		\begin{tabular}{|l|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} & \cellcolor{thread3} {\bf Thread 3} \\
			\hline
			\small \texttt{LOCK(mx);} & & \\
			\small \texttt{if~(!work\_exists())} & & \\
			\small \texttt{~~~~WAIT(cvar,~mx);} & & \\

			& \small \texttt{LOCK(mx);} & \\
			& \small \texttt{give\_work(work);} & \\
			& \small \texttt{SIGNAL(cvar);} & \\
			& \small \texttt{UNLOCK(mx);} & \\

			& & \small \texttt{LOCK(mx);} \\
			& & \small \texttt{work~=~claim\_work();} \\
			& & \small \texttt{UNLOCK(mx);} \\

			\small \texttt{work~=~claim\_work();~} & & \\
			\small \texttt{UNLOCK(mx);} & & \\
			\small \texttt{//~work~was~stolen} {\large \frownie}& & \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Testing}
	How can programmers be confident in their code's correctness?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in 15-410
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item Releasing to paying customers and worrying about correctness later
	\end{itemize}
	\linegap

	{\bf Motivation}: Can we do better than stress testing?
\end{frame}

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{largetest}, \texttt{mandelbrot} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	\linegap
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Systematic Testing}

\breakslide{\Large A different way of looking at race conditions\ldots}

\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork0.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork05.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork1.png}
	\end{center}
\end{frame}
\begin{frame}{Execution Tree}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{threadfork2.png}
	\end{center}
\end{frame}

% Say: Obviously we want to be able to always run the third path. But knowing
% that ahead of time is impossible.

% Say: In the rest of the lecture I'm going to refer to the "state space of
% possible interleavings", simply referring to this tree, and "size" of the state
% space refers to the number of branches.

\begin{frame}{Systematic Testing - The Big Picture}
	\textbf{Goal: Force the system to execute every possible interleaving.}
	% Say: checking for crashes/memory errors/etc in each one.
	\begin{itemize}
		\item On 1st execution, schedule threads arbitrarily until program ends.
		\item At end of each branch, rewind system and restart test.
		\item Artificially add preemptions to test different thread interleavings.
		\item Intuitively: Generate many ``tabular execution traces''.
	\end{itemize}
	\pause
	\linegap

	Okay, wait a sec...
	\begin{itemize}
		\item How can you possibly execute {\em every possible} interleaving?
		%\item How can you know where preempting matters most?
		\item How did you know to draw that tree's branches where they matter?
	\end{itemize}
\end{frame}

\begin{frame}{Preemption Points}
	\textbf{Preemption points} (PPs) are code locations where being preempted may cause different behaviour.
	\begin{itemize}
		\item IOW, somewhere that interesting interleavings can happen around.
	\end{itemize}
	\linegap

	Systematic tests are {\em parameterized} by the set of PPs.
	\begin{itemize}
		\item If there are $n$ PPs and $k$ threads, state space size is $n^k$.
		\item Need to choose the set of PPs very carefully for test to be effective.
	\end{itemize}
\end{frame}

\begin{frame}{Preemption Points}
	What does ``all possible interleavings'' actually mean?

	\linegap
	{\bf One extreme}: Preempt at {\em every instruction}
	\begin{itemize}
		\item Good news: Will find every possible race condition.
		\item Bad news: Runtime of test will be impossibly large.
	\end{itemize}
	\linegap

	{\bf Other extreme}: {\em Nothing} is a preemption point
	\begin{itemize}
		\item Good news: Test will finish quickly.
		\item Bad news: Only one execution was checked for bugginess.
		\begin{itemize}
			\item No alternative interleavings explored.
			\item Makes ``no race found'' a weak claim.
		\end{itemize}
	\end{itemize}
	\linegap

	Is there a {\bf ``sweet spot''}?
\end{frame}

\begin{frame}{Preemption Points}
	\textbf{Sweet spot}: Insert a thread switch everywhere it ``might matter''.

	\linegap
	When do we fear being preempted?
	\begin{itemize}
		\item Threads becoming runnable (\texttt{thr\_create()}, \texttt{cond\_signal()}, etc.)
			\begin{itemize}
				\item Preemptions may cause it to run before we're ready
			\end{itemize}
		\item Synchronization primitives (\texttt{mutex\_lock()}/\texttt{unlock()}, etc.)
			\begin{itemize}
				\item If buggy or used improperly\ldots
			\end{itemize}
		\item Unprotected shared memory accesses (``data races'')% Say: more on this later.
			\begin{itemize}
				\item May result in data structure corruption
			\end{itemize}
	\end{itemize}

	%%%% Not anymore! :D
	%Finding the sweet spot is a joint effort between programmer and tool. (More on this later.)
	%However, all these points together may still be too expensive...
\end{frame}

\begin{frame}{Challenges}
	Parameters of systematic tests must be kept small.
	\begin{itemize}
		\item Test program length / number of threads
		\item Number of preemption points used
	\end{itemize}
	\linegap

	\begin{center}
		\hspace{-0.25in}\includegraphics[width=0.3\textwidth]{expo.pdf} \\
		\vspace{-1.33in}\hspace{-1.75in}{\rotatebox{90}{\small state space size}} \\
		\vspace{0.15in}\hspace{-0.25in}{\small number of PPs}
	\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Landslide}

\breakslide{\Huge Landslide}

\begin{frame}{About The Project}
	\textbf{About me}: 4th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		\item 1st graduate year was 5th year MS, rest in Ph.D. program
		\begin{itemize}
			\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	{\bf About Landslide}
	\begin{itemize}
		\item Simics module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows that your {\em Pebbles} are less stable than you thought.
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Technical Overview}

\begin{frame}{Big Picture: Execution Tree Exploration}
	\textbf{Backtracking}
	\begin{itemize}
		\item At end of each branch, identify a PP to replay differently
		\item Reset machine state and start over
		\item Implemented using Simics bookmarks
			\begin{itemize}
				\item {\tt set-bookmark} and {\tt skip-to}
			\end{itemize}
		\item Replay test from the beginning, with a different interleaving
	\end{itemize}
	\pause
	\linegap

	{\bf Controlling scheduling decisions}
	\begin{itemize}
		\item Tool must control all sources of nondeterminism
		\item In 15-410, just timer and keyboard interrupts
			% optional - say: "in real kernels, disk/network io interrupts"
		\item Landslide repeatedly fires timer ticks until desired thread is run.
	\end{itemize}
\end{frame}

%\begin{frame}{Anatomy}
%	\begin{center}
%	\includegraphics[width=0.6\textwidth]{landslide.png}
%	\end{center}
%\end{frame}

\begin{frame}{Identifying Bugs}
	Landslide can {\em definitely discover}:
	\begin{itemize}
		\item Assertion failures % Say: note: the better your asserts are..!
		\item Segfaults
		\item Deadlock
		\item Use-after-free / double-free
	\end{itemize}
	\linegap
	Landslide can {\em reasonably suspect}:
	\begin{itemize}
		\item Infinite loop (halting problem)
		\item Data race bugs
	\end{itemize}
\end{frame}

\begin{frame}{What is a Data Race?}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item The threads can be reordered (e.g. no \texttt{cond\_signal()} in between)
	\end{itemize}
	\pause
	\linegap

	Data races are {\em not necessarily} bugs, just highly suspicious!
	\begin{itemize}
		\item Is unprotected {\tt foobar\_statistics\_counter++} bad?
		\item What about unprotected {\tt next\_thread\_id++}?
		\item ``If threads interleaved the wrong way here, it {\em might} crash later.''
		\begin{itemize}
			\item Hmmm...
		\end{itemize}
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Iterative Deepening}

\begin{frame}{Choosing the Right Preemption Points}
	How can we address exponential state space explosion?
	\pause
	\linegap

	State of the art tools hard-code a fixed set of preemption points.
	\begin{itemize}
		\item E.g., ``all thread library API calls'' or ``all kernel mutex locks/unlocks''
		\item Depending on length of test, completion time is unpredictable.
		\item More often, a subset is better in terms of time/coverage. 
	\end{itemize}
	\pause
	\linegap
	Current systematic testing model is not user-friendly.
	\begin{itemize}
		\item Tool: ``I want to use these PPs, but can't predict completion time.''
		\item User: ``I have 16 CPUs and 24 hours to test my program.''
	\end{itemize}
	\linegap
	{\em Stress testing offers this -- can we offer it too?}
\end{frame}

\begin{frame}{``Iterative Deepening'' of Preemption Points}
	Goal: Run the best tests for a given CPU budget.
	\linegap

	Based on experience from past 15-410 student volunteers
	\begin{itemize}
		\item Students worked best with an iterative process
		\item ``Start small, then add more preemption points as time allows''
		\item Landslide now automates this process
	\end{itemize}
	\linegap

	Named after analogous technique in chess AI.
	\begin{itemize}
		\item Chess search is DFS limited by max number of moves (ply).
		\item Chess AIs repeat DFS, increasing ply, until timeout.
	\end{itemize}
\end{frame}

\begin{frame}{Iterative Deepening}
	Repeat state space explorations, adding preemption points, until time is exhausted.
	\vspace{0.29in}
	\begin{center}
		\includegraphics[width=0.25\textwidth]{tree0.pdf}
	\end{center}
\end{frame}

\begin{frame}{Iterative Deepening}
	Repeat state space explorations, adding preemption points, until time is exhausted.
	\vspace{0.15in}
	\begin{center}
		\begin{tabular}{cc}
			\includegraphics[width=0.35\textwidth]{tree1.pdf} &
			\includegraphics[width=0.35\textwidth]{tree2.pdf}
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Iterative Deepening}
	Repeat state space explorations, adding preemption points, until time is exhausted.
	\begin{center}
		\includegraphics[width=0.5\textwidth]{tree3.pdf}
	\end{center}
	% say: and depending on your time budget, you might decide that you expect this one to be feasible, or you might fall back to a smaller state space from before.
\end{frame}

\begin{frame}{Iterative Deepening in Landslide}
	\textbf{Landslide-ID}: wrapper program for Landslide
	\begin{itemize}
		\item Manages work queue of jobs with different PPs
		\item Each job represents a new state space for Landslide to explore
		\item Prioritizes which jobs are important / likely to finish
		\begin{itemize}
			\item Based on nature of PPs (data races? mutexes?)
			\item Based on estimated completion time
		\end{itemize}
	\end{itemize}

	\linegap
	{\em Only required argument is CPU budget}
\end{frame}

\breakslide{\Huge Demo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Landslide for the People}

\begin{frame}{User Study}
	\textbf{Try Landslide on your P2!}
	\begin{itemize}
		\item Bare minimum effort: No more than 1 hour
		\begin{itemize}
			\item Clone a github URL, run setup script, run, answer survey questions
			\item Landslide will automatically report test results
		\end{itemize}
		\item Full study plan: 4-8 hours of active attention
		\begin{itemize}
			\item (Estimated, including time to diagnose and fix bugs)
			\item However, many tests should run passively overnight -- start soon!
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Prerequisites
	\begin{itemize}
		\item You {\em must} pass the P2 hurdle before using Landslide.
		\begin{itemize}
			\item {\tt startle}, {\tt agility\_drill}, {\tt cyclone}, {\tt join\_specific\_test}, {\tt thr\_exit\_join}
		\end{itemize}
		\item Must have {\em attempted} several stress tests
		\begin{itemize}
			\item {\tt juggle}, {\tt multitest}, {\tt racer} (for 1 hour), {\tt paraguay}
		\end{itemize}

	\end{itemize}
\end{frame}

\begin{frame}{User Study - Additional Information}
	Human Subjects Research
	\begin{itemize}
		\item CMU IRB has approved this study
		\item Landslide will collect results while you use it
		\begin{itemize}
			\item Record commands issued, take snapshots of your P2 code
			\item All data will be anonymized before publication
		\end{itemize}
		\item No coercion: {\bf \em There is no penalty for not participating.}
		\begin{itemize}
			\item I am not on course staff, cannot influence your grade
			\item Course staff will not have access to study data during semester
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Risks \& Benefits
	\begin{itemize}
		\item Benefit: Landslide may help you find/fix bugs, improving your grade!
		\item Risk: Landslide may find no bugs and be a waste of your time.
	\end{itemize}
	% Say: Accordingly, although I hope you will use landslide enough to get good results for me, you should also responsibly allocate time between testing with landslide versus actually implementing your P2.
\end{frame}

% IRB approved, plan on publishing collected data, anonymized
% Expect to spend X time
% Participation has no impact on grade but you will find bugs

\begin{frame}{User Study - How to Participate}
	To participate\dots
	\begin{itemize}
		\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Email me ({\tt bblum@cs.cmu.edu}) for instructions, survey
		\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	\linegap

	Watch your email for more details!
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{End}
\breakslide{\Large Questions?
\linegap

\begin{center}
	\includegraphics[width=0.65\textwidth]{3word-questions.png}
\end{center}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Bonus Slides}

\subsection{State Space Reduction}

\begin{frame}{Coping with State Space Explosion}
	Serious problem: State spaces grow exponentially
	\begin{itemize}
		\item With $p$ preemption points and $k$ runnable threads, size $p^k$.
		\item Threatens our ability to explore everything.
		\item Fortunately, some sequences result in identical states.
	\end{itemize}
	\linegap

	{\bf Partial Order Reduction} identifies and skips ``equivalent'' interleavings.
	\begin{itemize}
		\item After each execution, compare memory reads/writes of each thread.
		\item Find when reordering threads couldn't possibly change behaviour.
		\item Example follows\dots
	\end{itemize}
\end{frame}

\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond0.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}
\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{diamond1.png}
	\end{center}
\end{frame}

\subsection{Experimental Results}

\begin{frame}{Preliminary Evaluation}
	We tested 22 random submitted P2s from Spring 2014.
	\begin{itemize}
		\item Test cases:
		\begin{itemize}
			\item {\tt thr\_exit\_join}
			\item {\tt paraguay}
			\item {\tt rwlock\_downgrade\_read\_test}
			\item {\tt broadcast\_test} (new)
		\end{itemize}
		\item Each test given CPU budget of 11 cores $\times$ 10 minutes.
	\end{itemize}
\end{frame}

\begin{frame}{Preliminary Results}
	88 total tests were run (9,680 total CPU-minutes)
	\linegap

	Found 7 {\bf deterministic} initialization bugs (e.g. use-after-free)
	\linegap

	56 tests ran to completion without finding bugs:
	\begin{itemize}
		\item 47 exhausted the CPU budget.
		\item 9 completed all state spaces before time ran out.
	\end{itemize}
	\linegap

	Between 20 and 33 {\bf non-deterministic} bugs were found:
	\begin{itemize}
		\item 20 deadlock races among 12 P2s
		\item 10 use-after-free races among 6 P2s
		\item 3 misc (assert fail, segfault) among 2 P2s
	\end{itemize}
\end{frame}

\begin{frame}{More Preliminary Results}
	\textbf{Is dynamically adding ``data race'' PPs effective?}
	\begin{itemize}
		\item 3 among reported bugs (9-15\%) required data races to expose.
	\end{itemize}
	\linegap

	{\bf Is ``iterative deepening'' better than state-of-the-art?}
	\begin{itemize}
		\item Control experiment: Just 1 state space, same CPU time
		\begin{itemize}
			\item PPs used: \texttt{mutex\_lock()}, \texttt{mutex\_unlock()}
			\item 110 minutes on 1 CPU
		\end{itemize}
		\item Of 33 total bug reports, control failed to find 10 (30\%).
		\begin{itemize}
			\item 3 required data race PPs to expose
			\item 1 ran out of time
			\item 6 obscured by different bug in same state space
		\end{itemize}
	\end{itemize}
\end{frame}

\end{document}