### 2015-02-10 irb/hurdle.tex
Add 'hurdle'-like sign-up sheet for students to fill out.

\documentclass{article}
%\usepackage{amsmath,amsthm,amssymb,fullpage,yfonts,graphicx,proof,subfig,wrapfig,appendix,hyperref,mdwlist,wasysym}
\usepackage{amsmath,amsthm,amssymb,fullpage,yfonts,graphicx,proof,appendix,hyperref,mdwlist,wasysym}
\usepackage{upgreek}
%\usepackage{times}
\usepackage[charter]{mathdesign}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{epsfig}
\usepackage[bottom]{footmisc}
%\usepackage{mjz-titlepage}
\usepackage{framed}
\usepackage{setspace}
%\setstretch{1.05}
\usepackage{subfig}
\usepackage{changebar}
\usepackage{colortbl}
\usepackage{wrapfig}

\newcommand\true{\;\textit{true}}
\newcommand\false{\;\textit{false}}

\newcommand\alpher\alpha
\newcommand\beter\beta
\newcommand\gammer\gamma
\newcommand\delter\delta
\newcommand\zeter\zeta
\newcommand\Sigmer\Sigma

\newcommand\NN{\mathbb{N}}
\newcommand\QQ{\mathbb{Q}}
\newcommand\RR{\mathbb{R}}
\newcommand\ZZ{\mathbb{Z}}

\begin{document}
%\captionsetup{width=.75\textwidth,font=small,labelfont=bf}

\newcommand\checkbox[1]{\huge \Square & #1 \\}

\title{\bf Landslide: Systematic Testing in 15-410 \\ Sign-Up Sheet}
\author{Ben Blum}
\date{}
\maketitle

Thanks for volunteering to try out Landslide! To participate in the study, you and your partner must meet all the prerequisites below. Submit this form at my office (GHC 9009), any weekday between 13:00 and 17:00, or if you can't make it during those times, scan it after signing and email to {\tt bblum@cs.cmu.edu}.
\\

\begin{tabular}{lp{5.5in}}
	\checkbox{We have read the lecture slides, available on the 410 course website (or were present and paid attention in class on Wednesday).}

	\\
	\checkbox{We have read the study information sheet: \url{http://www.contrib.andrew.cmu.edu/~bblum/landslide-info-sheet.pdf}.}

	\\
	\checkbox{We agree that, upon receiving the instructions sheet, we will read it in its entirety before beginning any tests.}

	\\
	\checkbox{We pass the ``hurdle'' test suite listed in the P2 handout (on at least a vast majority of runs). This includes {\tt startle}, {\tt agility\_drill}, {\tt cyclone}, {\tt join\_specific\_test}, {\tt thr\_exit\_join}.}

	\\
	\checkbox{We have {\em attempted}, but not necessarily pass, the following stress tests; and {\em attempted} to debug and fix any observed bugs: {\tt juggle 4 3 2 0}, {\tt multitest}, {\tt racer} (for 15 minutes), {\tt paraguay}.}

\end{tabular}
\\
\vspace{0.5in}
\\
Group number: \underline{\hspace{1.5in}}\\
\vspace{0.5in}\\
Partner 1 andrewid: \underline{\hspace{1.5in}} Signature: \underline{\hspace{1.5in}} Date: \underline{\hspace{1in}}\\
\vspace{0.5in}\\
Partner 2 andrewid: \underline{\hspace{1.5in}} Signature: \underline{\hspace{1.5in}} Date: \underline{\hspace{1in}}

\end{document}