### 2016-04-30 proposal/410.tex
Add a bunch of boilerplate

\chapter{Education}
\label{chap:410}

This chapter proposes my plan to evaluate Landslide's effectiveness as a debugging aid for students in an educational setting.
This is the second of the three projects I am proposing for this thesis, and is currently ongoing work.