### 2016-04-30 proposal/lipservice.tex

\chapter{Broader Impact}
\label{chap:lipservice}

In this chapter I propose to add several enhancements to Landslide to allow it to be used beyond CMU's walls,
with little to no manual instrumentation effort required.
This is the third of the three projects I am proposing for this thesis and so far exists only in dreams.