### 2016-04-30 proposal/quicksand.tex

\chapter{State Spaces}
\label{chap:quicksand}

In this chapter, I will present a new algorithm for model checkers to automatically choose which preemption points to use to test reasonably-sized state spaces.
This is the first of the three projects I am proposing for this thesis, and has largely been completed and is under submission.