### 2016-05-13 proposal/related.tex
start writing related work

\chapter{Related Work}
\label{chap:related}

In this chapter I provide a brief tour of the related work.

\section{History of Stateless Model Checking}

{\bf Hall of Fame.}
Stateless model checking dates back to Verisoft \cite{verisoft}, the 1997 tool which first attempted to exhaustively explore the possibile ways to interleave threads.
Since then, researchers have built many tools along the same lines to test many kinds of programs.
%list of mcs follows
The most popular model checker, according to citation count, is Microsoft Research's CHESS \cite{chess},
a checker for userspace C++ programs which preempts at synchronization APIs.
%which introduced the Iterative Context Bounding technique \cite{chess-icb} for heuristically ordering the search to uncover bugs faster.
%
Other checkers include MaceMC \cite{macemc}, MoDist \cite{modist}, SAMC \cite{samc}, ETA \cite{dbug-retreat},
each of which limit thread communication to the boundaries of a message-passing API;
%
dBug \cite{dbug-ssv}, another CMU original, similar to CHESS;
%
and finally SPIN \cite{spin} and Inspect \cite{inspect} which can preempt at any shared variable access.
%
Perhaps by now, Landslide itself deserves a spot on this list: I first introduced it in my MS thesis \cite{landslide},
and it is most notable for testing kernel-level code through the use of a simulator.

{\bf Theoretical advances.}
% TODO

\section{Search Strategies}

% TODO
%etc..

\section{Educational Testing Tools}

\section{Real-World Techniques}