### 2017-01-04 proposal/410.tex
timeline

\chapter{Education}
\label{chap:410}

This chapter proposes my plan to evaluate Landslide's effectiveness as a debugging aid for students in an educational setting.
This is the second of the three projects I am proposing for this thesis, and is currently ongoing work.

\subsection{Landslide as a Debugging Tool}
\label{sec:studence}

\subsection{Landslide as a Grading Tool}
\label{sec:grading}

TODO