### 2017-09-27 irb/landslide-guide-p2.tex
add landslide guide p2

\documentclass{article}
%\usepackage{amsmath,amsthm,amssymb,fullpage,yfonts,graphicx,proof,subfig,wrapfig,appendix,hyperref,mdwlist,wasysym}
\usepackage{amsmath,amsthm,amssymb,fullpage,yfonts,graphicx,proof,appendix,hyperref,mdwlist,wasysym}
\usepackage{upgreek}
%\usepackage{times}
\usepackage[charter]{mathdesign}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{epsfig}
\usepackage[bottom]{footmisc}
%\usepackage{mjz-titlepage}
\usepackage{framed}
\usepackage{setspace}
%\setstretch{1.05}
\usepackage{subfig}
\usepackage{changebar}
\usepackage{colortbl}
\usepackage{wrapfig}

\newcommand\true{\;\textit{true}}
\newcommand\false{\;\textit{false}}

\newcommand\alpher\alpha
\newcommand\beter\beta
\newcommand\gammer\gamma
\newcommand\delter\delta
\newcommand\zeter\zeta
\newcommand\Sigmer\Sigma

\newcommand\NN{\mathbb{N}}
\newcommand\QQ{\mathbb{Q}}
\newcommand\RR{\mathbb{R}}
\newcommand\ZZ{\mathbb{Z}}

\begin{document}
%\captionsetup{width=.75\textwidth,font=small,labelfont=bf}

\title{\bf Landslide: Systematic Testing in 15-410 \\ User Guide}
\author{Ben Blum}
\date{}
\maketitle

\section{Introduction}

You should have received a document titled {\em User Study Information Sheet} that describes your rights as a study
participant and our confidentiality procedures. If you are missing this, please contact me immediately. My email
address is {\tt bblum@cs.cmu.edu} and my phone number is 412-304-4294; email is preferred.
\\

\noindent Likewise, please make sure you have read and understand the lecture slides on systematic testing, which are
available on the 15-410 course website. If you have any questions about the research background I would be
happy to answer them.

\section{Instructions}

We encourage you to use Landslide on the GHC cluster or UNIX pool machines ({\tt ghcXX.ghc.andrew.cmu.edu}
or {\tt unix.andrew.cmu.edu}). Landslide has not been tested with other configurations and will probably fail
mysteriously on your personal laptop, and we cannot promise to be of any help.

\begin{enumerate}
	\item Clone the repository - {\tt git clone https://github.com/bblum/landslide.git} - this will be your workspace.
	\item Run {\tt ./p2-setup.sh PATH}, where {\tt PATH} is the absolute path to your p2 implementation.
		\begin{itemize}
			\item This command will import your p2 into {\tt ./pebsim/p2-basecode/} and then attempt to build it.
				Whenever you update your p2 code, please do so in its original {\tt PATH}
				and then run {\tt p2-setup.sh} again.
			\item {\tt PATH} should be the path with {\tt 410kern}, {\tt 410user}, {\tt user/libthread}, etc as subdirectories, not the {\tt user/libthread} directory itself.
		\end{itemize}
	\item Run {\tt ./landslide OPTIONS} to run tests with Landslide, for various values of {\tt OPTIONS}:
		\begin{itemize}
			\item Use ./landslide -h to view command line options. The most important ones are -t , max testing
time, and -p , which test program to run.
			\item Landslide's bug reports will show up as HTML files.
				\begin{itemize}
					\item If you are using a cluster machine in person, you can view them by pointing a browser to \\
						{\tt file:///PATH/TO/landslide/FILENAME.html}
					\item For remote usage, move the html file to {\tt ~/www} in your AFS, and view it at \\
						{\tt http://www.contrib.andrew.cmu.edu/~YOUR\_ANDREWID/FILENAME.html}.
						Note the www in the URL is mandatory;
						note also that this will allow the world to see it,
						so be careful to clean up any traces in {\tt ~/www} when you're done!
				\end{itemize}
			\item Suggested test configurations to run are as follows.
				You can, of course, change the time limits or test programs as you see fit.
				For the longer-running configurations,
				we recommend running them inside of a {\tt screen} or {\tt tmux} session.
				\begin{itemize}
					\item \texttt{./landslide -t 30m -p thr\_exit\_join}
					\item \texttt{./landslide -t 30m -p mutex\_test}
					\item \texttt{./landslide -t 30m -p paradise\_lost}
					\item \texttt{./landslide -t 30m -p broadcast\_test}
					\item \texttt{./landslide -t 30m -p paraguay}
					\item \texttt{./landslide -t 30m -p rwlock\_downgrade\_read\_test}
						\\
					\item \texttt{./landslide -t 3h -p thr\_exit\_join}
					\item \texttt{./landslide -t 3h -p mutex\_test}
					\item \texttt{./landslide -t 3h -p paradise\_lost}
					\item \texttt{./landslide -t 3h -p broadcast\_test}
					\item \texttt{./landslide -t 3h -p paraguay}
					\item \texttt{./landslide -t 3h -p rwlock\_downgrade\_read\_test}
						\\
					\item \texttt{./landslide -t 12h -p thr\_exit\_join}
					\item \texttt{./landslide -t 12h -p mutex\_test}
					\item \texttt{./landslide -t 12h -p paradise\_lost}
					\item \texttt{./landslide -t 12h -p broadcast\_test}
					\item \texttt{./landslide -t 12h -p paraguay}
					\item \texttt{./landslide -t 12h -p rwlock\_downgrade\_read\_test}
				\end{itemize}
			\item To build your own custom test case for use with Landslide,
				put it in {\tt pebsim/p2-basecode/user/progs}
				and edit the {\tt STUDENTTESTS} line in {\tt pebsim/p2-basecode/config-incomplete.mk}
				(note: NOT {\tt config.mk}; the setup script auto-generates that file).
				Then run {\tt ./p2-setup.sh} again.
		\end{itemize}
	\item After fixing any found bugs, re-run Landslide with the same test options to confirm that your fix works.
\end{enumerate}

\noindent {\bf If something goes wrong:} If you find a Landslide bug, such as an assertion/crash, or such as a
bug report that you think is actually a bug in Landslide rather than a bug in your P2, or such as it getting stuck
and not exiting in the given time limit, please create a tarball of your workspace, just as you would to report
a reference kernel bug, but send it to Ben ({\tt bblum@cs.cmu.edu}) instead of 410 staff. For any other technical
issues getting it to work, or if you have difficulty understanding a bug report, send email to Ben.
\\

\noindent However, we cannot provide help on how to understand why your bug happened or how to fix it. The point of
the study is that Landslide's automation plus your own brain should be enough! On the other hand, if you have
a design question while deciding between potential ways to fix a bug, send it to 410 course staff as usual.

\end{document}