### 2018-10-06 thesis/thesis.tex
conclusion.

%for a more compact document, add the option openany to avoid
%starting all chapters on odd numbered pages
\documentclass[12pt]{cmuthesis}

% This is a template for a CMU thesis.  It is 18 pages without any content :-)
% The source for this is pulled from a variety of sources and people.
% Here's a partial list of people who may or may have not contributed:
%
%        bnoble   = Brian Noble
%        caruana  = Rich Caruana
%        colohan  = Chris Colohan
%        comar    = Cyrus Omar
%        jab      = Justin Boyan
%        josullvn = Joseph O'Sullivan
%        jrs      = Jonathan Shewchuk
%        kosak    = Corey Kosak
%        mjz      = Matt Zekauskas (mattz@cs)
%        pdinda   = Peter Dinda
%        pfr      = Patrick Riley
%        dkoes = David Koes (me)

% My main contribution is putting everything into a single class files and small
% template since I prefer this to some complicated sprawling directory tree with
% makefiles.

% some useful packages
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mdwlist}
\usepackage[numbers,sort]{natbib}
\PassOptionsToPackage{hyphens}{url}
\usepackage[backref,pageanchor=true,plainpages=false, pdfpagelabels, bookmarks,bookmarksnumbered,
%pdfborder=0 0 0,  %removes outlines around hyper links in online display
]{hyperref}
\usepackage{subfigure}
\usepackage{xspace}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{microtype}

%% custom fonts

%\usepackage{fontspec}
\usepackage{mathspec}
\usepackage{xeCJK}

%\setmainfont{XCharter}
\setallmainfonts(Digits,Latin){XCharter}
%\setmainfont{Utopia}[
%	Path = ./fonts/,
%	UprightFont = UtopiaStd-Regular.otf,
%	BoldFont = UtopiaStd-Semibold.otf,
%	ItalicFont = UtopiaStd-Italic.otf,
%	BoldItalicFont = UtopiaStd-SemiboldIt.otf
%	]

\setCJKmainfont{HeiseiMinStd-W5}[Path = ./fonts/]

% Approximately 1" margins, more space on binding side
%\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar]{geometry}
%for general printing (not binding)
\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar,hmarginratio=1:1]{geometry}

% Provides a draft mark at the top of the document.
\draftstamp{\today}{DRAFT}

% % nwf made this but i don't know how to get it to work?
% \usepackage{etoolbox}
% \makeatletter
% \apptocmd{\ref}{%
%   \immediate\write\@auxout%
%     {\expandafter\string\csname nwf@backref\endcsname{#1}{\arabic{page}}}%
% }{}{}
%
% \newcommand{\nwf@backref}[2]{%
%   \ifcsdef{nwf@backrefs@#1}%
%    {\csgappto{nwf@backrefs@#1}{, #2}}%
%    {\csgdef{nwf@backrefs@#1}{#2}}%
% }
% \AtEndDocument{\let\nwf@backref\relax}
%
% \newcommand{\linksto}[1]{\csname nwf@backrefs@#1\endcsname}
% \makeatother

\begin{document}
\frontmatter

%initialize page style, so contents come out right (see bot) -mjz
\pagestyle{empty}

\title{ %% {\it \huge Thesis Proposal}\\
{\bf Practical Concurrency Testing} \\
\normalsize \vspace{1em}
\begin{tabular}{c}
{\em or: How I Learned to Stop Worrying and Love the Exponential Explosion}
\end{tabular}}
\author{Ben Blum}
\date{Decembruary 2018}
\Year{2018}
\trnumber{}

\committee{
Garth Gibson, Chair \\
David A. Eckhardt \\
Brandon Lucia \\
Haryadi Gunawi, University of Chicago
}

\support{}
\disclaimer{}

% copyright notice generated automatically from Year and author.
% permission added if \permission{} given.

\keywords{landslide terminal, baggage claim, ground transportation, ticketing}

\maketitle

\begin{dedication}
For my family, my teachers, and my students.
\end{dedication}

\pagestyle{plain} % for toc, was empty

\begin{abstract}
Concurrent programming presents a challenge to students and experts alike because of the complexity of multithreaded interactions and the difficulty to reproduce and reason about bugs.
Stateless model checking is a concurrency testing approach which forces a program to interleave its threads in many different ways, checking for bugs each time.
This technique is powerful, in principle capable of finding any nondeterministic bug in finite time,
but suffers from exponential explosion as program size increases.
Checking an exponential number of thread interleavings is not a practical or predictable approach for programmers to find concurrency bugs before their project deadlines.

In this thesis, I develop several new techniques to make stateless model checking more practical for human use.
I have built Landslide, a stateless model checker specializing in undergraduate operating systems class projects.
Landslide extends the traditional model checking algorithm with a new framework for automatically managing
%the exploration of
multiple state spaces according to their estimated completion times,
which I show quickly finds bugs should they exist and also quickly verifies correctness otherwise.
I evaluate Landslide's suitability for inexpert use by presenting the results of many semesters providing it to students in 15-410, CMU's Operating System Design and Implementation class,
and more recently, students in similar classes at the University of Chicago and Penn State University.
%Finally, I will present several new techniques that allow stateless model checking to be practically employed on real-world programs.
%Finally, I will explore broader impact by extending Landslide to test some real-world programs and to be used by students at other universities.
Finally, I extend Landslide with a new concurrency model for hardware transactional memory,
and evaluate several real-world transactional benchmarks to show that stateless model checking can keep up with the developing concurrency demands of real-world programs.
\end{abstract}

\begin{acknowledgments}
I honor here the many people who supported me during these 7 years and made this journey possible.

\subsubsection{Thesis}

Professor Garth Gibson is largely responsible for shaping me into the researcher I am today.
%Thanks to Garth's guidance,
%this thesis has grown into something that past-me,
%first starting research with the simple vision ``make a better autograder for 410'',
%could not have then imagined.
Despite having primary research interests in other fields,
Garth enthusiastically took on the project,
and always pushed me to explore new problem domains,
to design scientifically thorough experiments,
to explain difficult concepts approachably,
to appreciate related work as charitably as possible,
and to seek guidance from industry people with relevant experience,
despite my often stubborn refusal to leave my little comfort zone.
So, even when it took me sending 10+ email reminders to get your attention,
truly,
thanks for everything, Garth.

Professor David Eckhardt
provided unending guidance on education, writing, career, life, you name it,
helped revise each semester's iteration of the recruiting lecture with a keen eye for the student mind,
contributed substantially to the design of the survey,
and supplied constant enthusiasm and encouragement that Landslide was a boon to 15-410.
I can't express my gratitude to Dave strongly enough,
even if it took until just this week for half of the Landslide-friendly tests to be included in the official P2 test suite.
%even though three of the Landslide-friendly tests never made it upstream into the official P2 test suite.

Ji\v{r}\'{i} \v{S}im\v{s}a, another student of Garth's,
first sparked my interest in the field,
and then gave invaluable support during my first few years.
Ji\v{r}\'{i}'s mentorship
oriented me in the world of research and helped me find my bearings as a grad student,
from teaching me how DPOR works
to helping revise my paper drafts
to offering the kind of candid career advice you can't get from a professor.
Had I not met Ji\v{r}\'{i} and learned about dBug during grad school applications way back,
none of this would have happened.

{Many others have collaborated with me directly on the research itself,
whom I thank here in approximate chapter order.
Michael J. Sullivan and Nathaniel Filardo were Landslide's first users besides myself back in its prototype days.
Joshua Wise, Miriam Zimmerman, and Glenn Willen helped revise my explanation of DPOR in Chapter~\ref{chap:background}.
Professor Brandon Lucia provided valuable comments and revisions to Chapter~\ref{chap:quicksand}'s content
that ultimately led to its publication at OOPSLA 2016.
Michael J. Sullivan double-checked the proofs therein.
Professor Timothy Zhu from Penn State University
and Professor Haryadi Gunawi and TA Kevin Zhao from the University of Chicago
graciously allowed me to use their students as guinea pigs
and helped greatly to improve Landslide's stability and robustness for use beyond CMU's walls,
greatly enriching Chapter~\ref{chap:education}'s evaluation.
David Blum and Skye Toor provided
invaluable advice on statistical analysis
that helped me evaluate the student
% justify last line bc of manual page break
\unskip\parfillskip 0pt \par}

% FIXME - cmon latex pls??
\pagebreak

\noindent
grade distributions therein.
Mario Dehesa-Azuara generously contributed the
transactional memory data structures
that make up about half of Chapter~\ref{chap:tm}'s evaluation suite.

Most importantly of all, I thank every student who ever used Landslide.
Though the introduction section may say differently,
their smiles and gratitude were the true motivation of this thesis.

\subsubsection{Academia}

The community of 15-410 TAs,
having surrounded me pretty much since starting undergrad,
nurtured my love of teaching
well before taking the class myself and well after graduation.
As skilled as the instructor is,
15-410 could not be the world's best operating systems class
without TAs who value thorough grading, the Socratic method,
and empathy to struggling students as much
from year to year as the crew I have known.
It was a blessing to serve, and later do research, among you all.
Thanks also to Wind River for generously providing CMU with Simics educational licenses.
I am also grateful to Adam Blank and Miriam Zimmerman
for founding and organizing the class Great Practical Ideas,
which teaches oft-neglected programming skills such as version control and debugging,
which I had the opportunity to teach as head instructor one semester.

I thank the Parallel Data Lab, my research group at CMU,
for supporting me at the PDL retreat every fall,
pushing me to refine my presentation skills and to network with strangers (now familiar faces) from industry.
I also thank those industry attendees from each of the PDL Consortium member companies
for their interest, even in more theoretical work,
and the enlightening conversations resulting therefrom.
I thank everyone outside my immediate research circle who,
whether at the retreat or otherwise,
insisted how cool my research was for directly helping students --
every such reminder helped me keep my head up during darker times.

I am grateful to the ThursDz Council,
my group of now-former grad student, Shadyside-bound (in spirit if not in person) restaurant-hoppers,
for both friendship and mentorship over the years.
In particular,
%in alphabetical for lack of a better order,
Chris Martens,
Jason Reed,
Jim McCann,
Rob Simmons,
Tom Murphy 7,
and
William Lovas:
you each have been role models to me
perhaps more than you know.

Without SIGBOVIK's unique brand of humorous, self-aware, and {\em definitely legitimate} research,
I probably would not have been interested in grad school in the first place,
or perhaps would have grown too frustrated along the way for lack of a cathartic outlet for conference paper woes.
I thank all who've worn the mantle of Harry Bovik over the years,
from the conference's original founders,
to every paper author,
to all future organizers who'll keep those ``mainstream'' conferences on their toes well into the future.

\subsubsection{Life}

% [ ] csd bgames
% [ ] undergrad frenz
% [ ] new frenz
% [ ] family, incl louie

Lastly,
the Android: Netrunner community has been such an important part of my life
for nearly as long as I've been a grad student.
Netrunner is a tabletop strategy deckbuilding card game
that depicts cyberspace hacking battles (completely different from real-life computer security)
alongside dystopian social issues (some of which seem oddly more familiar with each passing year).
Its characters have a diversity of ethnicity, gender identity, and sexual orientation
unparalleled in any other game I know of,
and the community that has grown around it reflects those values.
I thank in particular the other Stimhack moderators,
the members of {\sf \#rainbow-coalition},
and the local Pittsburgh crew
for their support and encouragement
and friendships very real and deep, even when forged only in text,
%\footnote{and Slack emojis},
that will long outlive the silly card game that brought us together.
%It'd be an injustice to list only a few names here and omit others, but you know who you are.

\end{acknowledgments}

\tableofcontents
% TODO: figure out how to make latex abbreviate these
\listoffigures
\listoftables

\mainmatter

%% Double space document for easy review:
%\renewcommand{\baselinestretch}{1.66}\normalsize

% The other requirements Catherine has:
%
%  - avoid large margins.  She wants the thesis to use fewer pages,
%    especially if it requires colour printing.
%
%  - The thesis should be formatted for double-sided printing.  This
%    means that all chapters, acknowledgements, table of contents, etc.
%    should start on odd numbered (right facing) pages.
%
%  - You need to use the department standard tech report title page.  I
%    have tried to ensure that the title page here conforms to this
%    standard.
%
%  - Use a nice serif font, such as Times Roman.  Sans serif looks bad.
%
% Other than that, just make it look good...

\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{orange}{RGB}{192,96,0}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,0,192}
% colors from sigbovik for tsx/etc
% threads
\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
\definecolor{seafoam}{RGB}{56,158,68} % * 2/3
\definecolor{salmon}{RGB}{208,89,74} % * 7/8
\definecolor{goldish}{RGB}{128,96,32}
% eval data
\definecolor{darkcyan}{RGB}{0,128,128}
\definecolor{lime}{RGB}{48,128,0}
\definecolor{darklavender}{RGB}{68,44,138}
\definecolor{brownish}{RGB}{128,96,0}
\definecolor{pinkish}{RGB}{128,34,102} % ff4488
% syntax hilighting
\newcommand\flow[1]{\hilight{brownish}{#1}\xspace}
\newcommand\const[1]{\hilight{brickred}{#1}\xspace}
\newcommand\call[1]{\hilight{pinkish}{#1}}
\newcommand\ctype[1]{\hilight{lime}{#1}\xspace}
\newcommand\ccomment[1]{\hilight{darkcyan}{#1}\xspace}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\newcommand\llama[1]{\includegraphics[width=#1]{llama.pdf}}
\newcommand\llitem{\item[\raisebox{-0.15em}{\llama{1.2em}}]}

\newcommand{\inspirationallinebreak}{\vspace{0.25em}}
\newcommand{\inspirationalhyphen}{-\hspace{-0.15em}-\xspace}
\newcommand\inspirationalquote[2]{\begin{flushright}
	\vspace{-1em}
	{
	\fontspec[Path=./fonts/]{AlexaStd}
	{\em #1}
	\inspirationallinebreak

	-\hspace{-0.15em}-\hspace{-0.15em}-{#2}
	}
	\vspace{2em}
\end{flushright}}

\newcommand\revision[1]{#1\xspace}

\input{intro}
\input{overview}
\input{landslide}
\input{quicksand}
\input{education}
\input{tm}
\input{related}
\input{future}

\chapter{Conclusion}
\label{chap:conclusion}
\inspirationalquote{
	\begin{tabular}{p{0.84\textwidth}}
	"If you are a god, Zeus, as the stories claim, then why did you create evolution?
	Why did you make a world that can only grow through cruelty and pain?"
	\\
	For a moment\inspirationalhyphen{}surely this meant death was near\inspirationalhyphen{}she thought she heard him answer:
	"My child, I was shaped by the gods that came before me, as you were shaped by me.
	The choice I had was between creation and oblivion, life and death.
	And I chose life, because any life is better than no life,
	because as long as there is life, there is hope\inspirationalhyphen{}if not for us, then for some generation to come."
	\\
	"Then how are you a god, if you can offer so little?" she whispered, feeling death creep closer.
	\\
	"I am a god because I take upon myself the burden of creation," the statue replied.
	\\
	"Then we are all gods," Alexandra said, and pushed the button.
	\end{tabular}}
%{Galatea, The Talos Principle}
{The Talos Principle}

Concurrency testing is, in a way, a microcosm for research in general:
an exponentially large problem
that invariably thwarts they whoever try to solve it ``perfectly'';
it bears fruit only to one who can find their own human-oriented meaning in partial results and compromises.
I speak of course literally on one side of this analogy, figuratively on the other.
This thesis has demonstrated such meaning for stateless model checking,
choosing perhaps one of the most challenging demographics of programmers in need of concurrency testing,
the undergraduate operating systems student,
as its target audience.
Along the way, it justified its techniques' applicability to problems beyond CMU's walls as well
by showing positive results at other universities
and by addressing new concurrency models developed by industry in just the last decade.
% also by comparing the preemption bounds of quicksand eval's bugs to prior work in ICB
% and finding them to actually be skewed to DEEPER bugs than in the "real" "world", so take that
As I write this I am also collaborating with 15-410 staff
to integrate Landslide as an official part of the course grading infrastructure,
so in some sense the na\"{i}ve wish to make an ``automatic TA'' that first made me interested in grad school
has also been fulfilled.

For as long as people have written science fiction books,
we have fantasized about robots that can make any complex intuitive decision a human can
in addition to the superhuman arithmetic ability that comes standard with silicon.
Recent trends in machine learning have pursued this aesthetic,
using statistics and pattern recognition to provide convincing output even on never-before-seen inputs,
but few such systems are designed to also alert the user when their output confidence is low.
% appropriate for serving ads (penalty for being wrong -- no click)
% appropriate for drug synthesis (clinical trial vets the output)
% not appropriate for... govt surveillance o.o;
% not appropriate for program verification
As tempting as it is to fantasize an oracle capable of telling for certain whether your program is free of concurrency bugs,
that would be a disservice to both user and machine:
the human skill that artificial intelligence should really try to emulate is
knowing its limits and asking for help when it gets confused.
% ways humans cooperate to make landslide better:
% - test case design incl. preemption sealing
% - abstraction reduction
% - manual inspection of data races

It's become fashionable among my peers lately,
whenever a new exploit is announced in the news,
to declare that security is doomed,
that computers were a mistake,
that we should all retire and become llama farmers,
and so on.
There is no doubt that we are burdened with massive technical debt from the mistakes made
during computer science's infancy.
Nevertheless,
safety-oriented programming languages and type systems are growing ever more popular,
and meanwhile formal verification techniques to check existing code,
written in the old unsafe ways, grow ever more powerful.
Landslide's ability to reach impressionable student minds is my little contribution in this long battle.
I have hope that we may one day live in peace with our silicon creations,
understanding and respecting their limitations just as they complement ours.
This thesis has been a profoundly meaningful journey for me and I hope you enjoyed reading it.

%\appendix
%\include{appendix}

\backmatter

%\renewcommand{\baselinestretch}{1.0}\normalsize

% By default \bibsection is \chapter*, but we really want this to show
% up in the table of contents and pdf bookmarks.
\renewcommand{\bibsection}{\chapter{\bibname}
\inspirationalquote{Only a doctor of philosophy, Darth.}{Robert Marsh}
}
%\renewcommand{\bibpreamble}{This text goes between the ``Bibliography''
%  header and the actual list of references}
\bibliographystyle{plainnat} % plain n'at
\bibliography{citations}

\end{document}